
<!-- rmarkdown v1 -->

Diversity Metrics Analysis Script
=================================

**K.M.Purcell**  
[email](mailto:kevin@kevin-purcell.com)


```{r prep, echo=FALSE,warning=FALSE,message=FALSE}
#load packages
library(reshape)
library(lattice)
library(MASS)
library(vegan)
library(doBy)
library(gplots)
library(segmented)
library(pander)
library(knitcitations)
```


```{r dataImport, echo=FALSE, warning=FALSE, message=FALSE}
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Data\\GatherSource\\importDiversityData.R")
```

## Data
### Fishery-independent

The datasets for this analysis derive from both fishery-dependent and fishery-indepedent data sets.  The fishery-independent data was derived from the SEAMAP survey data.  Data was downloaded on 2013-7-3, from the Gulf States Marine Fisheries online SEAMAP [access](http://seamap.gsmfc.org/) website.  We used several tables from their Access database with parameters including:


```{r}
names(starec.dat)
names(envrec.dat)
names(invrec.dat)
names(bgsrec.dat)
```

### Fishery-dependent
Data on fishery-dependent parameters (fishery effort) was derived from data from the SHRCOM databases, which I obtained via [Jim Nance](mailto:james.m.nance@noaa.gov) from the [Galveston Laboratory](http://www.galvestonlab.sefsc.noaa.gov/) of the [Southeast Fishery Science Center (SEFC)](http://www.sefsc.noaa.gov/).  Data from the fishery was in a aggregated format which was provided directly from Jim with the parameters:

```{r}
names(shrcom.agg)
```

Additionally, I utilized information obtained online from [Fishbase](http://www.fishbase.org/) for species description information, namely habitat characteristics vital to catagorizing species into groups.

```{r}
names(specinfo.dat)
```

The environmental drivers under consideration in this project were fishing effort, data obtained via the ```shrcom.agg``` above and coastal hypoxia.  The hypoxia data was obtained from Rabalais (XXXX)?  I did not update these values to reflect the new study ```citep("10.1021/es400983g").  

```{r}
summary(hypox.area)
```


```{r dataMerge, echo=FALSE,warning=FALSE,message=FALSE}
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Data\\GatherSource\\mergeDiversityData.R")
```




```{r regionalData, echo=FALSE,warning=FALSE,message=FALSE, cache=TRUE}
# Create seperate dataframes for Regions and seasons
sum.dat<-subset(complete.dat, complete.dat$month>=6&complete.dat$month<=8)
sum.la.dat<-subset(sum.dat, sum.dat$STAT_ZONE>=13&sum.dat$STAT_ZONE<=17)
sum.tx.dat<-subset(sum.dat, sum.dat$STAT_ZONE>=18&sum.dat$STAT_ZONE<=21)

fall.dat<-subset(complete.dat, complete.dat$month>=9&complete.dat$month<=11)
fall.la.dat<-subset(fall.dat, fall.dat$STAT_ZONE>=13&fall.dat$STAT_ZONE<=17)
fall.tx.dat<-subset(fall.dat, fall.dat$STAT_ZONE>=18&fall.dat$STAT_ZONE<=21)


## REFOMAT data.frames for diversity analysis
# Summer Louisiana Timeseries
sum.la.tab<-aggregate(ST_CNTEXP ~ YR + TAXONOMIC, data=sum.la.dat, FUN=sum)
sum.la.tab<-reshape(sum.la.tab, idvar="YR", timevar="TAXONOMIC", direction="wide")
sum.la.tab[is.na(sum.la.tab)]<-0
names(sum.la.tab) <- gsub("ST_CNTEXP.", "", names(sum.la.tab))  #removes the var name from reshape
YR<-sum.la.tab$YR
sum.la.tab$YR<-NULL

# Summer Texas Timeseries  
sum.tx.tab<-aggregate(ST_CNTEXP ~ YR + TAXONOMIC, data=sum.tx.dat, FUN=sum)
sum.tx.tab<-reshape(sum.tx.tab, idvar="YR", timevar="TAXONOMIC", direction="wide")
sum.tx.tab[is.na(sum.tx.tab)]<-0
names(sum.tx.tab) <- gsub("ST_CNTEXP.", "", names(sum.tx.tab))  #removes the var name from reshape
YR<-sum.tx.tab$YR
sum.tx.tab$YR<-NULL

# Fall Louisiana Timeseries
fall.la.tab<-aggregate(ST_CNTEXP ~ YR + TAXONOMIC, data=fall.la.dat, FUN=sum)
fall.la.tab<-reshape(fall.la.tab, idvar="YR", timevar="TAXONOMIC", direction="wide")
fall.la.tab[is.na(fall.la.tab)]<-0
names(fall.la.tab) <- gsub("ST_CNTEXP.", "", names(fall.la.tab))  #removes the var name from reshape
YR<-fall.la.tab$YR
fall.la.tab$YR<-NULL

# Fall Texas Timeseries  
fall.tx.tab<-aggregate(ST_CNTEXP ~ YR + TAXONOMIC, data=fall.tx.dat, FUN=sum)
fall.tx.tab<-reshape(fall.tx.tab, idvar="YR", timevar="TAXONOMIC", direction="wide")
fall.tx.tab[is.na(fall.tx.tab)]<-0
names(fall.tx.tab) <- gsub("ST_CNTEXP.", "", names(fall.tx.tab))  #removes the var name from reshape
YR<-fall.tx.tab$YR
fall.tx.tab$YR<-NULL
```

Data was then limited to the summer (```r min(sum.dat$month)``` - ```r max(sum.dat$month)```) and fall (```r min(fall.dat$month)``` - ```r max(fall.dat$month)```) months.  The nwGOM was divided into two regions Louisiana (```r min(fall.la.dat$STAT_ZONE)```- ```r max(fall.la.dat$STAT_ZONE)```) and Texas (```r min(fall.tx.dat$STAT_ZONE)```-```r max(fall.tx.dat$STAT_ZONE)```.  

A function ```divMetrics()``` was written to calcuate all the diversity metrics of interest for each of the region and seasonal data sets and returen a data frame of values.


```{r divData, echo=TRUE,warning=FALSE,message=FALSE}
# load diversity function
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\functions\\divMetrics.R")

# calculate diversity metrics
sum.la.agg<-divMetrics(sum.la.tab)
sum.la.agg$mod<-"sum.la"

sum.tx.agg<-divMetrics(sum.tx.tab)
sum.tx.agg$mod<-"sum.tx"

fall.la.agg<-divMetrics(fall.la.tab)
fall.la.agg$mod<-"fall.la"

fall.tx.agg<-divMetrics(fall.tx.tab)
fall.tx.agg$mod<-"fall.tx"
```

```{r plots, echo=FALSE, warning=FALSE, error=FALSE, fig.align='center'}
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Analysis\\ResultsFigures\\LaDivMetricsFig.R")
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Analysis\\ResultsFigures\\TxDivMetricsFig.R")
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Analysis\\ResultsFigures\\LaDivSmoothFig.R")
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Analysis\\ResultsFigures\\TxDivSmoothFig.R")
```


Data frames were created for each of the region/seasonal data sets.


```{r lowessCalcs, echo=FALSE, message=FALSE, warning=FALSE} 
### Piecewise regression
# Create a df for each lowess smooth
sum.la.S<-as.data.frame(lowess(sum.la.agg$S~sum.la.agg$YR));sum.la.S$mod<-"sum.la"
sum.la.d<-as.data.frame(lowess(sum.la.agg$d~sum.la.agg$YR));sum.la.d$mod<-"sum.la"
sum.la.J<-as.data.frame(lowess(sum.la.agg$J~sum.la.agg$YR));sum.la.J$mod<-"sum.la"
sum.la.shan<-as.data.frame(lowess(sum.la.agg$shan~sum.la.agg$YR));sum.la.shan$mod<-"sum.la"
sum.la.simp<-as.data.frame(lowess(sum.la.agg$simp~sum.la.agg$YR));sum.la.simp$mod<-"sum.la"
sum.la.N1<-as.data.frame(lowess(sum.la.agg$N1~sum.la.agg$YR));sum.la.N1$mod<-"sum.la"
sum.la.N2<-as.data.frame(lowess(sum.la.agg$N2~sum.la.agg$YR));sum.la.N2$mod<-"sum.la"
 
sum.tx.S<-as.data.frame(lowess(sum.tx.agg$S~sum.tx.agg$YR));sum.tx.S$mod<-"sum.tx"
sum.tx.d<-as.data.frame(lowess(sum.tx.agg$d~sum.tx.agg$YR));sum.tx.d$mod<-"sum.tx"
sum.tx.J<-as.data.frame(lowess(sum.tx.agg$J~sum.tx.agg$YR));sum.tx.J$mod<-"sum.tx"
sum.tx.shan<-as.data.frame(lowess(sum.tx.agg$shan~sum.tx.agg$YR));sum.tx.shan$mod<-"sum.tx"
sum.tx.simp<-as.data.frame(lowess(sum.tx.agg$simp~sum.tx.agg$YR));sum.tx.simp$mod<-"sum.tx"
sum.tx.N1<-as.data.frame(lowess(sum.tx.agg$N1~sum.tx.agg$YR));sum.tx.S$N1<-"sum.tx"
sum.tx.N2<-as.data.frame(lowess(sum.tx.agg$N2~sum.tx.agg$YR));sum.tx.S$N2<-"sum.tx"

fall.la.S<-as.data.frame(lowess(fall.la.agg$S~fall.la.agg$YR));fall.la.S$mod<-"fall.la"
fall.la.d<-as.data.frame(lowess(fall.la.agg$d~fall.la.agg$YR));fall.la.d$mod<-"fall.la"
fall.la.J<-as.data.frame(lowess(fall.la.agg$J~fall.la.agg$YR));fall.la.J$mod<-"fall.la"
fall.la.shan<-as.data.frame(lowess(fall.la.agg$shan~fall.la.agg$YR));fall.la.shan$mod<-"fall.la"
fall.la.simp<-as.data.frame(lowess(fall.la.agg$simp~fall.la.agg$YR));fall.la.simp$mod<-"fall.la"
fall.la.N1<-as.data.frame(lowess(fall.la.agg$N1~fall.la.agg$YR));fall.la.N1$mod<-"fall.la"
fall.la.N2<-as.data.frame(lowess(fall.la.agg$N2~fall.la.agg$YR));fall.la.N2$mod<-"fall.la"

fall.tx.S<-as.data.frame(lowess(fall.tx.agg$S~fall.tx.agg$YR));fall.tx.S$mod<-"fall.tx"
fall.tx.d<-as.data.frame(lowess(fall.tx.agg$d~fall.tx.agg$YR));fall.tx.d$mod<-"fall.tx"
fall.tx.J<-as.data.frame(lowess(fall.tx.agg$J~fall.tx.agg$YR));fall.tx.J$mod<-"fall.tx"
fall.tx.shan<-as.data.frame(lowess(fall.tx.agg$shan~fall.tx.agg$YR));fall.tx.shan$mod<-"fall.tx"
fall.tx.simp<-as.data.frame(lowess(fall.tx.agg$simp~fall.tx.agg$YR));fall.tx.simp$mod<-"fall.tx"
fall.tx.N1<-as.data.frame(lowess(fall.tx.agg$N1~fall.tx.agg$YR));fall.tx.N1$mod<-"fall.tx"
fall.tx.N2<-as.data.frame(lowess(fall.tx.agg$N2~fall.tx.agg$YR));fall.tx.N2$mod<-"fall.tx"


#plot
#par(mfrow=c(2,4),pty="s")
#la summer
par(mfrow=c(2,2))
plot(sum.la.agg2$S.norm.mean~sum.la.agg2$YR, pch="",
     main="Louisiana Summer",
     xlab="Year",
     ylab="Normalized value",
     ylim=c(0.454,1.1))
#points(sum.la.agg2$N.norm.mean~sum.la.agg2$YR, type="p", pch=2)
lines(lowess(sum.la.dat2$S.norm~sum.la.dat2$YR), lty=1)
lines(lowess(sum.la.dat2$N.norm~sum.la.dat2$YR), lty=2)
lines(lowess(sum.la.dat2$d.norm~sum.la.dat2$YR), lty=3)
lines(lowess(sum.la.dat2$J.norm~sum.la.dat2$YR), lty=4)
lines(lowess(sum.la.dat2$shan.norm~sum.la.dat2$YR), lty=5)
lines(lowess(sum.la.dat2$simp.norm~sum.la.dat2$YR), lty=6)
lines(lowess(sum.la.dat2$N1.norm~sum.la.dat2$YR), lty=7)
lines(lowess(sum.la.dat2$N2.norm~sum.la.dat2$YR), lty=8)
abline(v=2003,col=4,lty=2) #effort
abline(v=1996,col=9,lty=2) #hypoxia
rect(1998,0,2003,2, density=20, col="grey",
     angle=-30, border="transparent") 

#la fall
plot(fall.la.agg2$S.norm.mean~fall.la.agg2$YR, pch="",
     main="Louisiana fall",
     xlab="Year",
     ylab="Normalized value",
     ylim=c(0.45,1.2))
#points(fall.la.agg2$N.norm.mean~fall.la.agg2$YR, type="p", pch=2)
lines(lowess(fall.la.dat2$S.norm~fall.la.dat2$YR), lty=1)
lines(lowess(fall.la.dat2$N.norm~fall.la.dat2$YR), lty=2)
lines(lowess(fall.la.dat2$d.norm~fall.la.dat2$YR), lty=3)
lines(lowess(fall.la.dat2$J.norm~fall.la.dat2$YR), lty=4)
lines(lowess(fall.la.dat2$shan.norm~fall.la.dat2$YR), lty=5)
lines(lowess(fall.la.dat2$simp.norm~fall.la.dat2$YR), lty=6)
lines(lowess(fall.la.dat2$N1.norm~fall.la.dat2$YR), lty=7)
lines(lowess(fall.la.dat2$N2.norm~fall.la.dat2$YR), lty=8)
abline(v=2003,col=4,lty=2)
abline(v=1996,col=9,lty=2) #hypoxia
rect(1998,0,2003,2, density=20, col="grey",
     angle=-30, border="transparent") 

#tx summer
plot(sum.tx.agg2$S.norm.mean~sum.tx.agg2$YR, pch="",
     main="Texas Summer",
     xlab="Year",
     ylab="Normalized value",
     ylim=c(0.454,1.1))
#points(sum.tx.agg2$N.norm.mean~sum.tx.agg2$YR, type="p", pch=2)
lines(lowess(sum.tx.dat2$S.norm~sum.tx.dat2$YR), lty=1)
lines(lowess(sum.tx.dat2$N.norm~sum.tx.dat2$YR), lty=2)
lines(lowess(sum.tx.dat2$d.norm~sum.tx.dat2$YR), lty=3)
lines(lowess(sum.tx.dat2$J.norm~sum.tx.dat2$YR), lty=4)
lines(lowess(sum.tx.dat2$shan.norm~sum.tx.dat2$YR), lty=5)
lines(lowess(sum.tx.dat2$simp.norm~sum.tx.dat2$YR), lty=6)
lines(lowess(sum.tx.dat2$N1.norm~sum.tx.dat2$YR), lty=7)
lines(lowess(sum.tx.dat2$N2.norm~sum.tx.dat2$YR), lty=8)
abline(v=2003,col=4,lty=2)
abline(v=1996,col=9,lty=2) #hypoxia
rect(1999,0,2003,2, density=20, col="grey",
     angle=-30, border="transparent") 

#tx fall
plot(fall.tx.agg2$S.norm.mean~fall.tx.agg2$YR, pch="",
     main="Texas fall",
     xlab="Year",
     ylab="Normalized value",
     ylim=c(0.45,1.2))
#points(fall.tx.agg2$N.norm.mean~fall.tx.agg2$YR, type="p", pch=2)
lines(lowess(fall.tx.dat2$S.norm~fall.tx.dat2$YR), lty=1)
lines(lowess(fall.tx.dat2$N.norm~fall.tx.dat2$YR), lty=2)
lines(lowess(fall.tx.dat2$d.norm~fall.tx.dat2$YR), lty=3)
lines(lowess(fall.tx.dat2$J.norm~fall.tx.dat2$YR), lty=4)
lines(lowess(fall.tx.dat2$shan.norm~fall.tx.dat2$YR), lty=5)
lines(lowess(fall.tx.dat2$simp.norm~fall.tx.dat2$YR), lty=6)
lines(lowess(fall.tx.dat2$N1.norm~fall.tx.dat2$YR), lty=7)
lines(lowess(fall.tx.dat2$N2.norm~fall.tx.dat2$YR), lty=8)
abline(v=2003,col=4,lty=2)
abline(v=1996,col=9,lty=2) #hypoxia
rect(1998,0,2005,2, density=20, col="grey",
     angle=-30, border="transparent") 
```



```{r fig1, echo=FALSE,warning=FALSE,message=FALSE, fig.align='center'}

shrcom.agg2<-aggregate(effort~ yr, data=shrcom.agg, FUN=sum)
shrcom.agg2<-subset(shrcom.agg2, shrcom.agg2$yr>=1987)

shrcom<-data.frame(shrcom.agg2$effort, shrcom.agg2$yr)
colnames(shrcom)[1]<-'y'
colnames(shrcom)[2]<-'x'

biomass.dat<-aggregate(cpue~YR, data=complete.dat, FUN=mean)
abundance.dat<-aggregate(catch_cpue~YR, data=complete.dat, FUN=mean)

biomass.sm<-as.data.frame(lowess(biomass.dat$cpue~biomass.dat$YR))



source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\functions\\oldBreakpt.R")


oldBreakpt(biomass.dat,1,2)
oldBreakpt(abundance.dat,1,2)
oldBreakpt(shrcom.agg2,1,2)
oldBreakpt(hypox.area,1,2)


# source to figure 1 code
# 4 panel plot of Abundance-Biomass time-series and Hypoxia and Effort time-series
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\Analysis\\ResultsFigures\\figure1.R")

```

```{r, fig.align='center', fig.height=6, fig.width=8}
# load the break point function
source("C:\\Users\\Kevin.Purcell\\Documents\\GitHub\\EcosysMetricsGOM\\functions\\breakpts.R")
breakpt.fun(sum.la.S,1,2,1,1)
breakpt.fun(sum.la.d,1,2,2,1) 
breakpt.fun(sum.la.J,1,2,3,1) 
breakpt.fun(sum.la.shan,1,2,4,1) 
breakpt.fun(sum.la.simp,1,2,5,1) 
breakpt.fun(sum.la.N1,1,2,6,1) 
breakpt.fun(sum.la.N2,1,2,7,1) 

breakpt.fun(sum.tx.S,1,2,1,2)
breakpt.fun(sum.tx.d,1,2,2,2) 
breakpt.fun(sum.tx.J,1,2,3,2) 
breakpt.fun(sum.tx.shan,1,2,4,2) 
breakpt.fun(sum.tx.simp,1,2,5,2) 
breakpt.fun(sum.tx.N1,1,2,6,2) 
breakpt.fun(sum.tx.N2,1,2,7,2) 

breakpt.fun(fall.la.S,1,2,1,3)
breakpt.fun(fall.la.d,1,2,2,3) 
breakpt.fun(fall.la.J,1,2,3,3) 
breakpt.fun(fall.la.shan,1,2,4,3) 
breakpt.fun(fall.la.simp,1,2,5,3) 
breakpt.fun(fall.la.N1,1,2,6,3) 
breakpt.fun(fall.la.N2,1,2,7,3) 

breakpt.fun(fall.tx.S,1,2,1,4)
breakpt.fun(fall.tx.d,1,2,2,4) 
breakpt.fun(fall.tx.J,1,2,3,4) 
breakpt.fun(fall.tx.shan,1,2,4,4) 
breakpt.fun(fall.tx.simp,1,2,5,4) 
breakpt.fun(fall.tx.N1,1,2,6,4) 
breakpt.fun(fall.tx.N2,1,2,7,4) 

breakpt.tab<-rbind(sum.la_S, sum.la_d, sum.la_J, sum.la_shan, sum.la_simp, sum.la_N1, sum.la_N2,
                   sum.tx_S, sum.tx_d, sum.tx_J, sum.tx_shan, sum.tx_simp, sum.tx_N1, sum.tx_N2,
                   fall.la_S, fall.la_d, fall.la_J, fall.la_shan, fall.la_simp, fall.la_N1, fall.la_N2,
                   fall.tx_S, fall.tx_d, fall.tx_J, fall.tx_shan, fall.tx_simp, fall.tx_N1, fall.tx_N2)
#bind the model output tables
rm(sum.la_S, sum.la_d, sum.la_J, sum.la_shan, sum.la_simp, sum.la_N1, sum.la_N2,
                   sum.tx_S, sum.tx_d, sum.tx_J, sum.tx_shan, sum.tx_simp, sum.tx_N1, sum.tx_N2,
                   fall.la_S, fall.la_d, fall.la_J, fall.la_shan, fall.la_simp, fall.la_N1, fall.la_N2,
                   fall.tx_S, fall.tx_d, fall.tx_J, fall.tx_shan, fall.tx_simp, fall.tx_N1, fall.tx_N2)
```

```{r brkptTable, results='asis'}
pandoc.table(breakpt.tab)
```


 











