---
title: "Size-Based Indicator Analysis"
author: "Kevin M. Purcell"
date: "Tuesday, November 18, 2014"
output: html_document
---

## Objectives

Two primary sources for metric obtions and preliniary metric response are [Shin et al. 2005](http://icesjms.oxfordjournals.org/content/62/3/384) and [Rochet and Trinkle 2003](http://www.nrcresearchpress.com/doi/abs/10.1139/f02-164).  


## Methods

1. Bring in size frequency data
2. Attached locations or sampling sites to size based data
3. Derive size based indicators from initial size sampling data
4. Evaluate temporal patterns of indicators


```{r dataImport, echo=FALSE}
sfDat <- read.table("C:/Users/Kevin.Purcell/Documents/seamap_2013/GLFREC.csv", 
                    header = TRUE, sep = ",")
```



```{r varTable, echo=FALSE, results='asis', warning=FALSE, error=FALSE}

shortnames=names(sfDat)
#create variable descriptions
description <- c("unique integer (int) entry id",
                 "unique int for ea. entry in CRUSIES table",
                 "Unique int for ea. entry in BGSREC table",
                 "unique int for ea. entry in STAREC table",
                 "unique int for ea. entry in VESSELS table",
                 "four character (ch.) string format YYXX",
                 "five ch. string for Pascagoula Station No.",
                 "9 digit field for Genus species biocode",
                 "no current used",
                 "7 ch. field for Genus",
                 "6 ch. field for species",
                 "4 digit XXX.X field for weight (kg)",
                 "2 ch. field containing measurement code",
                 "4 ch. field (numeric) of length (mm)",
                 "1 ch. field representing sex (M, F, U) can be blank",
                 "1 ch. field representing maturity code")

metadataBound <- cbind(shortnames, description)
colnames(metadataBound) <- c("Parameter", "Description")
metadataTable <- xtable::xtable(metadataBound, 
                                align="ccl",
                                caption="Data variables and description") 
Metadata <- xtable::print.xtable(metadataTable, type="html")
```


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
