---
title: "Size-Based Indicator Analysis"
author: "Kevin M. Purcell"
date: "Tuesday, November 18, 2014"
output: html_document
---

## Objectives
Identify and calculate size-based indicator (SBI) time series for the Gulf of Mexico based on the length-frequency data collected by [SEAMAP]() for our focal species.  Indicatory identification is derived from two primary sources [Shin et al. 2005](http://icesjms.oxfordjournals.org/content/62/3/384) and [Rochet and Trinkle 2003](http://www.nrcresearchpress.com/doi/abs/10.1139/f02-164).  Initially a global time series will be developed for community-wide metrics of mean length:

$$\bar{L}=\frac{\sum\limits_{N}L}{N}$$  


will also be generated and trends over the survey period will be assessed.




## Methods

```{r envSetup, echo=FALSE, warning=FALSE, message=FALSE}
#Clear the junk
graphics.off()
rm(list=ls(all=TRUE))


#library(magrittr)
library(dplyr)
library(stargazer)
library(ggplot2)
```




```{r dataImport, echo=FALSE}
# Import Size frequency table from SEAMAP_2013 DB
# sfDat <- read.csv("C:/Users/Kevin.Purcell/Documents/seamap_2013/GLFREC.csv", 
#                    header = TRUE, sep = ",")

sfDat <- read.csv("C:/Users/Kevin.Purcell/Documents/seamap_2013/GLFREC2.csv", 
                    header = TRUE, sep = ",")

# GLFREC2.csv was the second attempt to extract the data from the Access DB.
# The initial export (ie GLFREC.csv) only exported around 1M values as opposed
# the ~3M in the database.

# Making a measurement type factor to make sorting easier
# create a new variable in df
sfDat$meas_typ <- sfDat$MEASCD_GLF

# Recode the new variable to letter designations for filtering
sfDat$meas_typ[sfDat$meas_typ==1] <- "FL"
sfDat$meas_typ[sfDat$meas_typ==2] <- "SL"
sfDat$meas_typ[sfDat$meas_typ==3] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==4] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==5] <- "WD"
sfDat$meas_typ[sfDat$meas_typ==6] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==8] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==10] <- "WD"
sfDat$meas_typ[sfDat$meas_typ==11] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==12] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==13] <- "ML"
sfDat$meas_typ[sfDat$meas_typ==14] <- "WD"
sfDat$meas_typ[sfDat$meas_typ==15] <- "TRD"
sfDat$meas_typ[sfDat$meas_typ==16] <- "WD"
sfDat$meas_typ[sfDat$meas_typ==17] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==18] <- "TL"
sfDat$meas_typ[sfDat$meas_typ==19] <- "UNKNOWN"
sfDat$meas_typ[sfDat$meas_typ==20] <- "OTHER"
sfDat$meas_typ[sfDat$meas_typ==22] <- "WD"
sfDat$meas_typ[sfDat$meas_typ==23] <- "SA"


# Rename BIO_GLF to BIO_CODE
# BIO_CODE will become the standard for df merging
colnames(sfDat)[8] <- "BIO_CODE"


# Import station table from SEAMAP_2013 DB
# This data is primarily for data info by station
stDat <- read.csv("C:/Users/Kevin.Purcell/Documents/seamap_2013/STAREC.csv", 
                    header = TRUE, sep = ",")

# Create a yr and month variable from MO_DAY_YR field
stDat$date<-stDat$MO_DAY_YR
stDat$date<-as.Date(stDat$date, "%m/%d/%Y")
stDat$month<-format(stDat$date, format="%m")
stDat$YR<-format(stDat$date, format="%Y")
stDat$YR<-as.numeric(stDat$YR)
stDat$month<-as.numeric(stDat$month)

# Limit df to necessary values
stDat <- dplyr::select(stDat, STATIONID, month, YR)



# Import the Taxonomy data from SEAMAP_2013 DB
#**** Commented out due to better data in "info.dat" file ****#

taxDat <- read.csv("C:/Users/Kevin.Purcell/Documents/seamap_2013/NEWBIOCODESBIG.csv", header = TRUE, sep = ",")

# This file had info on a much larger number of species from the GOM
# specinfo.dat<-read.csv("C:/Users/kevin.purcell/Documents/comm_analysis/spec_info_summary.csv")
# colnames(specinfo.dat)[1]<-'BIO_BGS'


# Import the study species list with taxonomy, TL, Habitat, and Bycatch designations
# Primiarily favored this because it is only our 65 study species and includes bycatch designations
info.dat <- read.csv("C:/Users/kevin.purcell/Documents/comm_analysis/matrix_plot/infoDat.csv")


# Import BGSREC table for the station by species specific count data
bgsDat <- read.csv("C:/Users/Kevin.Purcell/Documents/seamap_2013/BGSREC.csv", header = TRUE, sep = ",")

# Filter only needed columns
# names(bgsDat)
# head(bgsDat)

# bgsDat1 <- select(bgsDat, STATIONID, CNTEXP, BIO_BGS, IS_SAMPLE)
# head(sbiDat3)


#Import list of study species
studySpecies<-read.csv("C:/Users/kevin.purcell/Documents/comm_analysis/seamap_keepers_comm_analysis.csv")
studySpecies$keep <- NULL
colnames(studySpecies)[1] <- "BIO_GLF"
# lenCount <- sbiDat3 %>%
#   filter(STATIONID==4) %>%
#   group_by(BIO_GLF, common_name) %>%
#   summarize(spcCnt=n(), meanL)
```



```{r dataTabl1, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='asis'}

shortnames=names(sfDat)
#create variable descriptions
description <- c("unique integer (int) entry id",
                 "unique int for ea. entry in CRUSIES table",
                 "Unique int for ea. entry in BGSREC table",
                 "unique int for ea. entry in STAREC table",
                 "unique int for ea. entry in VESSELS table",
                 "four character (ch.) string format YYXX",
                 "five ch. string for Pascagoula Station No.",
                 "9 digit field for Genus species biocode",
                 "no current used",
                 "7 ch. field for Genus",
                 "6 ch. field for species",
                 "4 digit XXX.X field for weight (kg)",
                 "2 ch. field containing measurement code",
                 "4 ch. field (numeric) of length (mm)",
                 "1 ch. field representing sex (M, F, U) can be blank",
                 "1 ch. field representing maturity code")

metadataBound <- cbind(shortnames, description)
colnames(metadataBound) <- c("Parameter", "Description")
stargazer(metadataBound, type = "html", summary = FALSE)

# Do this later to avoid choking on data

#sbiDat1 <- dplyr::inner_join(sfDat, stDat, by = "STATIONID", copy = FALSE)




#return all rows from x where there are matching values in y, keeping just columns from x. 
# Essentially filter data to only Species in our "LIST"
# sbiDat2 <- dplyr::inner_join(sbiDat1, studySpecies, by = NULL, copy = FALSE)
# names(sbiDat2)
# # check
# length(unique(sbiDat2$BIO_GLF))



#change the column name in taxDat to BIO_GLF
colnames(taxDat)[3] <- "BIO_GLF"

#join taxonomy data to database
#sbiDat3 <- dplyr::inner_join(sbiDat2, taxDat, by = NULL, copy = FALSE)
```




```{r dataTabl2, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='asis'}
# species table
specTable <- dplyr::inner_join(studySpecies, taxDat, by = NULL, copy = FALSE) %>%
  select(BIO_GLF, TAXONOMIC, common_name)
colnames(specTable)[1] <- "Code"
colnames(specTable)[2] <- "Genus species"
colnames(specTable)[3] <- "Common name"

stargazer(specTable, type = "html", summary = FALSE, digit.separator = "")
```


```{r dataMung, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}

# Rename column for merging
#head(info.dat)
colnames(info.dat)[2] <- "BIO_CODE"

# First create a data set and calculations for non-weighted mean length 
# based on the entire commmunity
# First calculate the number measured and sum of the lengths
# for each species at each station

# Limit species data to only neccessary 
info.dat2 <- select(info.dat, BIO_CODE, common_name, bycatch)
names(info.dat2)


head(bgsDat)
colnames(bgsDat)[15] <- "BIO_CODE"
bgsDat2 <- select(bgsDat, STATIONID, BIO_CODE, CNTEXP)

head(sfDat)
sfDat <- filter(sfDat, meas_typ=="TL")
sfDat2 <- select(sfDat, STATIONID, BIO_CODE, LEN_GLF)
sfDat2 <- inner_join(sfDat2, stDat, by= "STATIONID", copy=FALSE)

# SBI data for the entire GOM
SBIDAT.ALLSPEC <- inner_join(sfDat2, info.dat2, by = "BIO_CODE" , copy = FALSE)
SBIDAT.ALLSPEC <- inner_join(SBIDAT.ALLSPEC, bgsDat2, by = c("STATIONID","BIO_CODE") , copy = FALSE)

SBIDAT.ALLSPEC <- filter(SBIDAT.ALLSPEC, LEN_GLF>=0 & LEN_GLF<300)
SBIDAT.ALLSPEC$wLen <- SBIDAT.ALLSPEC$LEN_GLF*SBIDAT.ALLSPEC$CNTEXP
SBIDAT.ALLSPEC <- filter(SBIDAT.ALLSPEC, wLen>0)
```

```{r dataPlot, echo=FALSE, warning=FALSE, error=FALSE, fig.align='center'}

SBIDAT.ALLSPEC2 <- SBIDAT.ALLSPEC %>%
  group_by(YR) %>%
  summarize(wLensum=sum(wLen), cntsum=sum(CNTEXP))

SBIDAT.ALLSPEC2$WmeanL <- SBIDAT.ALLSPEC2$wLensum/SBIDAT.ALLSPEC2$cntsum

# Plot of non-weighted trends
ggplot(data=SBIDAT.ALLSPEC2, aes(x=YR, y=WmeanL)) + 
  geom_line() + 
  geom_point() + 
  theme_bw()


# Look at trends by bycatch designation group

SBIDAT.ALLSPEC3 <- SBIDAT.ALLSPEC %>%
  group_by(YR, bycatch) %>%
  summarize(wLensum=sum(wLen), cntsum=sum(CNTEXP))
SBIDAT.ALLSPEC3$WmeanL <- SBIDAT.ALLSPEC3$wLensum/SBIDAT.ALLSPEC3$cntsum


ggplot(data=SBIDAT.ALLSPEC3, aes(x=YR, y=WmeanL, group=bycatch, colour=bycatch)) + 
  geom_line() + 
  geom_point() + 
  theme_bw()

```





